Improved Cation Binding to Lipid Bilayer with Negatively Charged POPS by Effective Inclusion of Electronic Polarization

### 2019-10-18 Manuscript/manuscript.tex
Small edits in simulation details

\subsection{Measurements of acyl chain order parameters}
A R-type Proton Detected Local Field (R-PDLF) experiment was performed to determine C--H bond order
parameters
\begin{equation}\label{OPequation}
  S_{\rm{CH}}=\frac{1}{2}\langle3\cos^2\theta-1\rangle,
\end{equation}
  where $\theta$ denotes the angle of the C--H bond
with the bilayer normal and the angular brackets define a time average on a time scale of approximately 1 microsecond.
The experiment was done using a Bruker Avance III 400 spectrometer operating at a $^1$H Larmor frequency of 400.03 MHz
equipped with a standard 4~mm CP-MAS HXY probe. The set up of the R-PDLF experiment was the following
(using the notation from figures 1c and 2c in the original publication describing the R-PDLF experiment \cite{dvinskikh04}).
The magic angle spinning (MAS) frequency used was 5.15 kHz. The recoupling pulses in the R18 blocks had
therefore a length of $10.79 \, \mu \mathrm{s}$, which correspond to a nutation frequency of 46.35~kHz. Increments of $18 \times 2$ recoupling
pulses where used giving a spectral width of 2.6~kHz in the indirect dimension and a total number of 32 points in
the indirect dimension were recorded. These settings enable to record dipolar slices with a C--H bond order parameter
resolution of $\pm$0.01. The C--H bond order parameter determined from a given dipolar
splitting, $\Delta\nu$ (see e.g. Fig. \ref{R-PDLFslices} in supplementary information) is equal to $\Delta\nu/(0.315\times21.5 kHz)$, where 0.315 is
the scaling factor of the R18 recoupling sequence and 21.5 kHz is the maximum $^1$H-$^{13}$C dipolar coupling for a C--H bond.
The refocused-INEPT transfer \cite{morris79,burum80} 
was used for transferring the polarization from $^1$H nuclei to the covalently bond $^{13}$C nuclei with the delays of $\tau_1 = 1.94$~ms and $\tau_2 = 0.97$~ms
(set as multiples of the MAS rotation period). The pulses for the refocused-INEPT sequence had a nutation frequency equal to 63.45~kHz.
For acquiring the spectra, a total number of 1024 transients were recorded for each point in the indirect dimension, using an acquisition
time of 0.1~s under SPINAL64 $^1$H decoupling \cite{fung00} with a nutation frequency of 50~kHz, and with dwell time giving a spectral
width of 200 ppm for the $^{13}$C chemical shift dimension. The set of free induction decays recorded were then processed as
described e.g. in Ref.~\citenum{ferreira08}. The dipolar splittings in the crowded spectral region
between 29 and 30 ppm were assigned based on the previous assignments for POPC \cite{ferreira13} 
and the POPS order parameters calculated from simulations in this work.

@@

In the ECC-lipids simulations,
the ECC-POPC parameters~\cite{melcr18} (available from Ref. \citenum{ECC-POPC_nacl_cacl2_files}), 
the ECC-POPS parameteres derived in this work (see above for details and Ref. \citenum{ecclipids_pcps_nacl_kcl_series} for parameters), 
the SPC/E~\cite{Berendsen1987} water model, and ECC-ions \cite{martinek17, kohagen16, Pluharova2014} were used.
The SPC/E water model was selected because its lower dielectric constant is consistent with the
ECC concept \cite{leontyev11,leontyev14}.

The C--H bond order parameters were calculated directly from Eq.~\ref{OPequation}.
The time average was first calculated for each lipid, and
the standard error of the mean from these individual values was then used as the
error estimate \cite{botan15,ollila16,NMRlipidsIV}.
Python program that uses the
MDAnalysis library \cite{agrawal11,gowers16} is available in Ref.~\citenum{MATCHgit} ({\tt scripts/calcOrderParameters.py}). 
The ion number density profiles were calculated using the {\tt gmx density} tool
of the Gromacs sofware package \cite{gromacsMANUAL}.
The salt concentrations in buffer before solvating the lipids, reported in the used experimental data set \cite{roux90},
were calculated as $[{\rm salt}]=N_{\rm c} \times[{\rm water}]\,/\,N_{\rm w}$,
where $N_{\rm c}$ is the number of cations in simulation, $[{\rm water}]$\,=\,55.5~M and $N_{\rm w}$
is the number of water molecules in the simulation.
As discussed in our previous work \cite{NMRlipidsIV}, the hydration levels of
multilamellae are expected to be sufficiently similar in the used simulations and reference experiments \cite{roux90}.

>>>>

\subsection{Measurements of acyl chain order parameters}
A R-type Proton Detected Local Field (R-PDLF) experiment was performed to determine C--H bond order
parameters
\begin{equation}\label{OPequation}
  S_{\rm{CH}}=\frac{1}{2}\langle3\cos^2\theta-1\rangle,
\end{equation}
  where $\theta$ denotes the angle of the C--H bond
with the bilayer normal and the angular brackets define a time average on a time scale of approximately 1 microsecond.
The experiment was done using a Bruker Avance III 400 spectrometer operating at a $^1$H Larmor frequency of 400.03 MHz
equipped with a standard 4~mm CP-MAS HXY probe. The set up of the R-PDLF experiment was the following
(using the notation from figures 1c and 2c in the original publication describing the R-PDLF experiment \cite{dvinskikh04}).
The magic angle spinning (MAS) frequency used was 5.15 kHz. The recoupling pulses in the R18 blocks had
therefore a length of $10.79 \, \mu \mathrm{s}$, which correspond to a nutation frequency of 46.35~kHz. Increments of $18 \times 2$ recoupling
pulses where used giving a spectral width of 2.6~kHz in the indirect dimension and a total number of 32 points in
the indirect dimension were recorded. These settings enable to record dipolar slices with a C--H bond order parameter
resolution of $\pm$0.01. The C--H bond order parameter determined from a given dipolar
splitting, $\Delta\nu$ (see e.g. Fig.~\ref{R-PDLFslices} in supplementary information) is equal to $\Delta\nu/(0.315\times21.5 kHz)$, where 0.315 is
the scaling factor of the R18 recoupling sequence and 21.5 kHz is the maximum $^1$H-$^{13}$C dipolar coupling for a C--H bond.
The refocused-INEPT transfer \cite{morris79,burum80} 
was used for transferring the polarization from $^1$H nuclei to the covalently bond $^{13}$C nuclei with the delays of $\tau_1 = 1.94$~ms and $\tau_2 = 0.97$~ms
(set as multiples of the MAS rotation period). The pulses for the refocused-INEPT sequence had a nutation frequency equal to 63.45~kHz.
For acquiring the spectra, a total number of 1024 transients were recorded for each point in the indirect dimension, using an acquisition
time of 0.1~s under SPINAL64 $^1$H decoupling \cite{fung00} with a nutation frequency of 50~kHz, and with dwell time giving a spectral
width of 200 ppm for the $^{13}$C chemical shift dimension. The set of free induction decays recorded were then processed as
described e.g. in Ref.~\citenum{ferreira08}. The dipolar splittings in the crowded spectral region
between 29 and 30 ppm were assigned based on the previous assignments for POPC \cite{ferreira13} 
and the POPS order parameters calculated from simulations in this work.

@@

In the ECC-lipids simulations, we used
the ECC-POPC parameters~\cite{melcr18} (available from Ref. \citenum{ECC-POPC_nacl_cacl2_files}), 
the ECC-POPS parameteres derived in this work (see above for details and Ref. \citenum{ecclipids_pcps_nacl_kcl_series} for parameters), 
the SPC/E~\cite{Berendsen1987} water model, and ECC-ions \cite{martinek17, kohagen16, Pluharova2014}. 
The SPC/E water model was selected because its lower dielectric constant is consistent with the
ECC concept \cite{leontyev11,leontyev14}.

The C--H bond order parameters were calculated directly from Eq.~\ref{OPequation}
using a Python code openly available at Ref.~\citenum{MATCHgit} ({\tt scripts/calcOrderParameters.py}) 
that uses the MDAnalysis library \cite{agrawal11,gowers16}.  
The time average was first calculated for each lipid, and
the standard error of the mean from these individual values was then used as the
error estimate \cite{botan15,ollila16,NMRlipidsIV}.
The ion number density profiles were calculated using the {\tt gmx density} tool
of the Gromacs sofware package \cite{gromacsMANUAL}.
In simulations, the salt concentrations in a buffer before solvating the lipids, 
as reported in the used experimental data set \cite{roux90},
were calculated as $[{\rm salt}]=N_{\rm c} \times[{\rm water}]\,/\,N_{\rm w}$,
where $N_{\rm c}$ is the number of additional cations, 
$N_{\rm w}$ is the number of water molecules,
and $[{\rm water}]$\,=\,55.5~M. 
As discussed in our previous work \cite{NMRlipidsIV}, the hydration levels of
multilamellae are expected to be sufficiently similar in the used simulations and reference experiments \cite{roux90}.
